The effect of authoritarian legislature on regime-opposition interaction

### 2014-11-22 writeups/experiment_checking.tex
done with all, doing references

Finally, to guard against systematic bias in non-response, I check the co-variate balance between firms that responded and those that did not in \autopageref{tab:balance_response}. The results suggest some issues with co-variate balance: firms with educated CEOs and CEOs with former ties to state businesses respond slightly more. This means that the population observed in the experiment may be different, leading to difficulty in generalizing our ATE (average treatment effect) to PATE (population average treatment effect). However, educated businessmen with political ties are precisely the group most likely to be co-opted, this is in fact the group that we should pay most attention to. Therefore, while our treatment effect may not generalize to ordinary Vietnamese businessman, it does comport with the focus of our study on political elites.

>>>>

Finally, to guard against systematic bias in non-response, I check the co-variate balance between firms that responded and those that did not in \autoref{tab:balance_response}. The results suggest some issues with co-variate balance: firms with educated CEOs and CEOs with former ties to state businesses respond slightly more. This means that the population observed in the experiment may be different, leading to difficulty in generalizing our ATE (average treatment effect) to PATE (population average treatment effect). However, educated businessmen with political ties are precisely the group most likely to be co-opted, this is in fact the group that we should pay most attention to. Therefore, while our treatment effect may not generalize to ordinary Vietnamese businessman, it does comport with the focus of our study on political elites.

### 2014-11-22 writeups/experiment_description.tex
writing on survey result
done with all, doing references

From the large-n analysis, we find suggestive evidence of an authoritarian legislature having a moderating effect on political violence. However, while the matching design alleviates our concern about functional misspecification and incorrect counterfactual, the threat of selection on unobservable remains. Indeed, large variation in the country and country-year intercepts suggests that our model of violence is not yet precise.

@@

The survey experiment is as follows. Conceptually, the treatment is a country ``having an elected legislature,'' which cannot be randomized. However, in our survey design we can prompt the respondents to think about the legislature as a bargaining arena. My survey question is included in Vietnam's annual business survey PCI (Provincial Competitiveness Index) as follows:

@@

Given this mismatch between political and economic power, Vietnam is an insightful case study of the co-optation theory with the new class of capitalists being the potential opposition. Indeed, after the 1992 Constitutional revision, the Vietnamese National Assembly (VNA) has included more diverse and assertive, amending law proposed by the executive, albeit without rejecting it. The representation of businessmen and women in the National Assembly has also increased. In 2007 only 1 out of 30 self-nominated candidates won seat in the VNA, while in 2011 4 out of 15 candidates won seats, including two of the country's best known capitalists. (cite http://blogs.reuters.com/global/2011/06/04/party-wins-big-in-vietnam/)

>>>>

From the large-n analysis, we find suggestive evidence of an authoritarian legislature having a modest effect on political violence. However, while the matching design alleviates our concern about functional misspecification and incorrect counterfactual, the threat of selection on unobservable remains. Indeed, large variation in the country and country-year intercepts suggests that our model of violence is not yet precise.

@@

The survey experiment is as follows. Conceptually, the treatment is a country ``having an elected legislature,'' which cannot be randomized. However, in our survey design we can prompt the respondents to think about the legislature as a bargaining arena. My survey question is included in Vietnam's annual business survey PCI (Provincial Competitiveness Index) as follows:\footnote{The PCI is the largest annual business survey in Vietnam, which is designed to assess the ease of doing business, economic governance, and administrative reform efforts \citep{Malesky2013}. This year the surveys reaches 8,013 firms. The PCI is already well-known and implemented regularly, raising no concern that the business may respond differently if they are aware of being studied (i.e. the Hawthorne effect).}

@@

Given this mismatch between political and economic power, Vietnam is an insightful case study of the co-optation theory with the new class of capitalists being the potential opposition. Indeed, after the 1992 Constitutional revision, the Vietnamese National Assembly (VNA) has included more diverse and assertive, amending law proposed by the executive, albeit without rejecting it. The representation of businessmen and women in the National Assembly has also increased. In 2007 only 1 out of 30 self-nominated candidates won seat in the VNA, while in 2011 4 out of 15 candidates won seats, including two of the country's best known capitalists \citep{Ruwitch2011}.

### 2014-11-22 writeups/experiment_result.tex
done with all, doing references

\begin{table}[!htbp] \centering 
  \caption{Treatment effect in Survey experiment} 
  \label{tab:survey_result} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & Contacting the National Assembly is effective \\ 
\hline \\[-1.8ex] 
 Treatment prompt & 0.264$^{***}$ \\ 
  & (0.073) \\ 
  & \\ 
 Constant & 0.929$^{***}$ \\ 
  & (0.049) \\ 
  & \\ 
\hline \\[-1.8ex] 
Observations & 3,951 \\ 
Log Likelihood & $-$2,253.137 \\ 
Akaike Inf. Crit. & 4,510.274 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

>>>>

The size of this treatment effect is substantively significant given that the presence of businessmen in Vietnam's National Assembly is still limited. Furthermore, policy making in Vietnam is still dominated by the Executive branch, with the National Assembly mainly deliberating scripted agenda prepared by the Standing Committee \citep{Montesano2005}. Despite these limitations on the participation of business in policy making through the National Assembly, we see that business are still willing to participate in the legislature. Therefore, the treatment effect in this survey experiment is likely to be even larger in more open authoritarian legislatures.

\begin{table}[H] 
  \centering 
  \caption{Treatment effect in Survey experiment} 
  \label{tab:survey_result} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & Contacting the National Assembly is effective \\ 
\hline \\[-1.8ex] 
 Treatment prompt & 0.264$^{***}$ \\ 
  & (0.073) \\ 
  & \\ 
 Constant & 0.929$^{***}$ \\ 
  & (0.049) \\ 
  & \\ 
\hline \\[-1.8ex] 
Observations & 3,951 \\ 
Log Likelihood & $-$2,253.137 \\ 
Akaike Inf. Crit. & 4,510.274 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

### 2014-11-22 writeups/introduction.tex
done with all, doing references

\section{Introduction}

Once disregarded as mere ``window-dressing'', authoritarian institutions have garnered the attention of scholars as an important tool for power sharing, which, in turn, leads to various substantive effect, including regime survival, increased investment, better economic performance.

While our understanding of authoritarian institution has branched to numerous substantive topics, its causal root deserves more careful studies. Given its common definition as the ``rule of the game'', institutions are fundamentally a set of constraints of human interactions, and thus its effect must be most evident in the interaction between the strategic actors itself.  When scholars study the effect of institutions on various substantive areas, inevitably the causal chain must pass through the strategic actions of the actors, mostly through the logic of credible and regular spoil sharing (cite Przworski, Svolik). This causal step is theoretically crucial to all of our findings regarding authoritarian institutions – the goal of this article is, therefore, to empirically examine it.

Beyond its theoretical contribution, a study of violence in partially liberalized dictatorship has immense real-world importance. Indeed, violence during political liberalization is the greatest concern of many would-be democrats living under authoritarian regimes. For example, Chinese citizens show overwhelming enthusiasm when asked about democratic features such as popular accountability, separation of power, and political liberalism, yet remain consistently cautious of instability caused by transition (Chu 2008, 309). Understanding whether and why countries experience increased violence during liberalization is crucial for the prospect of democracy.

\section{Argument and Structure}

This article argues that if the authoritarian legislature is indeed a credible commitment and informational device, then we should see decreased level of violence between the regime and the dissdent in the presence of an authoritarian legislature. Indeed, taking a page from the rationalist war literature, I explains how political violence should only happen when the regime and the opposition misjudge one another's capacity and preference. Given transparent information and a commitment device, regardless of the capacity and preference of both sides, negotiating to reach an agreement is always preferable to violence due to the savings in the social cost of violence.

Armed with this theoretical argument, I empirically examine the level of violence in regime-dissident interaction using a novel event dataset across authoritarian regime post Cold War. I constructed the data by looking explicitly at countries that create their legislature during their authoritarian spells and matched these countries with those of the same conditions but did not create legislature. In addition to this large-n investigation, I zoom in at the causal mechanism with a survey experiment implemented in Vietnam, a hard case for the theory of legislature on regime-dissident interaction. These studies show that blah blah blah.

The article will proceed as follows:

>>>>

\section{Introduction}
\label{sec:introduction}

Once disregarded as mere ``window-dressing,'' authoritarian institutions have garnered the attention of scholars as an important tool for power sharing, which, in turn, leads to various substantive effects, including regime survival, increased investment, and better economic performance \citep{Gandhi2008}.

While our understanding of authoritarian institution has branched to numerous substantive topics, its causal root deserves more careful studies. Given its common definition as the ``rule of the game'', institutions are fundamentally a set of constraints on human interactions, and thus its effect must be most evident in the interaction between strategic actors itself.  When scholars study the effect of institutions on various substantive areas, inevitably the causal chain must pass through the strategic actions of the actors, mainly with the logic of credible and regular spoil sharing (Svolik, Przeworski). This causal step is theoretically crucial to all of our findings regarding authoritarian institutions. The goal of this article is, therefore, to empirically examine it.

Beyond its theoretical contribution, a study of violence in partially liberalized dictatorship has immense real-world importance. Indeed, violence during periods of political liberalization is the greatest concern of many would-be democrats living under authoritarian regimes. For example, Chinese citizens show overwhelming enthusiasm when asked about democratic features such as popular accountability, separation of power, and political liberalism, yet remain consistently cautious of instability caused by transition (Chu 2008, 309). Therefore, understanding the pattern of violence during liberalization is crucial for the prospect of democracy.

The article will proceed as follows. \Cref{sec:argument} presents an overview of the theoretical argument and empirical analyses. \Cref{sec:theory} posits two theoretical explanations for the effect of legislature on violence. \Cref{sec:empirics_largeN} uses a Bayesian hierarchical model to analyze political violence in authoritarian regimes post-Cold War, showing that the legislature has a modest positive effect. \Cref{sec:empirics_survey} uses a survey experiment to demonstrate that co-optation strategy does prompt Vietnamese business to interact with the regime, a pre-requisite for any material outcome such as violence to change. \Cref{sec:conclusion} concludes.

### 2014-11-22 writeups/multilevel_empiricalstrategy.tex
done with all, doing references

\section{Empirics}

\subsection{Description of data}

\begin{itemize}

\item Dependent variable: \textit{legislature}, operationalized as $LIEC \geq 5$

As my theoretical argument emphasizes the role of the legislature as an informational and commitment device, I code a country-year as having a legislature if multiple parties are legal. Even if opposition parties do not win seat, the electoral process also conveys information about the organizational strength and the preference of both sides. This coding rule operationalizes as having LIEC (Legislative Index of Electoral Competitiveness) be 5 or higher in the DPI (Database of Political Institutions).\footnote{The values LIEC are as follows: 4---1 party, multiple candidates; 5---multiple parties are legal but only one party won seats; 6---multiple parties did win seats but the largest party received more than 75\% of the seats; 7---largest party got less than 75\% (DPI 14)}

\item Independent variable: \textit{violence} in regime-dissident interaction

Using a dataset of machine coded news reports around the world from 1991-2012, I am able to look at individual events that happened between the regime and the dissident, enabling an analysis at a highly granular level. These event data come from ICEWS (Integrated Crisis Early Warning System), a project funded by DARPA (Defense Advanced Research Projects Agency) to aid the US policy makers in predicting and responding to crises around the world . These news articles are harvested from ``over 75 international sources (AP, UPI, and BBC Monitor) as well as regional sources (India Today, Jakarta Post, Pakistan Newswire, and Saigon Times),'' ensuring comprehensive coverage  (cite O'Brien 2010 (94), 2013). While this dataset has been de-duplicated once so that an event covered by multiple news report only shows up once, I created my own filter to delete duplicates that share the same source actor, target actor, and happen at the same time and place.

These text data are then processed with Schrodt's TABARI event data coding system to determine the nature and the intensity of the reported political activity. In addition, the event is matched with a dictionary of actors (with over 8,000 entries) that further divides into five broad categories: religious, governmental, dissidents, business, and other. In sum, from this database I am able to extract any interaction between the government and the dissident, where and when it happens, and its level of violence (measured by the Goldstein score\footnote{The Goldstein score is a measure of intensity of conflict or cooperation, developed by Joshua S. Goldstein. The score has values from -10 (Military attack, clash, assault), 0 (Explain or state policy, state future position) to 8.3(Extend military assistance) See Goldstein 375 for how the event types are coded into this scale).}
\end{itemize}

@@

This strategy is implemented as follows. First, to construct the treated population, I look at all country years in which an authoritarian legislature is created and build a four-year panel around it, including the one year before the treatment, and the three years after. Second, to construct the control population, I build similar four-year panels from all the other authoritarian country years without a legislature. Finally, for each panel in my treated population, I find its closest match from the control population based on their first year (which is the pre-treatment year for the treated countries). I match on level of violence and other covariates as reported in \autoref{tab:matching_balance_largen}, significantly improving the balance across the board. The matched pairs of panel are reported in \autoref{tab:matched_pairs}. Given that factors that affect violence may have various non-linear effect, matching alleviates the serious concern about functional misspecification.

@@

\begin{sidewaystable}[ht]
\centering
\begin{tabular}{lrrlrrrl}
  \hline
 & \multicolumn{2}{c}{Pre-Match} &   \multicolumn{2}{c}{Post-Match} \\
Covariate & Treated mean & Control mean & p-value & Treated mean & Control mean & p-value &  \\ 
  \hline
Ethnic Fractionalization \_ & 0.61 & 0.59 & 0.59 & 0.61 & 0.60 & 0.81 &  \\ 
Violence (Goldstein score) & -4.50 & -4.05 & 0.51 & -4.50 & -4.49 & 0.94 &  \\ 
$\text{Violence}^2$ & 29.84 & 23.22 & 0.34 & 29.84 & 29.03 & 0.65 &  \\ 
Regime duration (years) & 12.08 & 34.30 & 0.00*** & 12.08 & 12.71 & 0.71 &  \\ 
$\text{Regime duration}^2$ & 278.00 & 2783.57 & 0.00** & 278.00 & 287.38 & 0.89 &  \\ 
Military & 0.08 & 0.06 & 0.67 & 0.08 & 0.12 & 0.57 &  \\ 
Monarchy & 0.04 & 0.25 & 0.00*** & 0.04 & 0.04 & 1.00 &  \\ 
Party & 0.29 & 0.48 & 0.07* & 0.29 & 0.29 & 1.00 &  \\ 
log(gdp) & 22.65 & 23.95 & 0.00*** & 22.65 & 22.78 & 0.67 &  \\ 
log(gdp per capita) & 7.63 & 8.58 & 0.00*** & 7.63 & 7.65 & 0.95 &  \\ 
Military expenditure (\% of GDP) & 2.44 & 4.34 & 0.00*** & 2.44 & 2.40 & 0.89 &  \\
Land (1000 $\text{km}^2$)& 760.22 & 1674.68 & 0.00*** & 760.21 & 783.37 & 0.80 &  \\ 
Population (million) & 22.62 & 140.73 & 0.00*** & 22.62 & 21.74 & 0.90 &  \\
Natural Resource Export (\% of GDP) & 14.79 & 22.64 & 0.03** & 14.79 & 14.16 & 0.72 &  \\ 
Resource $\times$ Ethnic & 8.61 & 14.43 & 0.01** & 8.61 & 8.33 & 0.82 &  \\ 
   \hline
\end{tabular}
\caption{Pre and Post-Matching Balance}
\label{tab:matching_balance_largen}
\end{sidewaystable}

>>>>

@@

This strategy is implemented as follows. First, to construct the treated population, I look at all country years in which an authoritarian legislature is created and build a four-year panel around it, including the one year before the treatment, and the three years after. Second, to construct the control population, I build similar four-year panels from all the other authoritarian country years without a legislature. Finally, for each panel in my treated population, I use genetic matching to find its closest match from the control population based on their first year (which is the pre-treatment year for the treated countries). I match on level of violence and other covariates as reported in \autoref{tab:matching_balance_largen}, significantly improving the balance across the board. The matched pairs of panel are reported in \autoref{tab:matched_pairs}. Given that factors that affect violence may have various non-linear effect, matching alleviates the serious concern about functional misspecification.

@@

\begin{sidewaystable}[ht]
\centering
\begin{tabular}{lrrlrrrl}
  \hline
 & \multicolumn{2}{c}{Pre-Match} &   \multicolumn{2}{c}{Post-Match} \\
Covariate & Treated & Control & p-value & Treated & Control & p-value &  \\ 
  \hline
Ethnic Fractionalization & 0.61 & 0.59 & 0.59 & 0.61 & 0.60 & 0.81 &  \\ 
Violence (Goldstein score) & -4.50 & -4.05 & 0.51 & -4.50 & -4.49 & 0.94 &  \\ 
$\text{Violence}^2$ & 29.84 & 23.22 & 0.34 & 29.84 & 29.03 & 0.65 &  \\ 
Regime duration (years) & 12.08 & 34.30 & 0.00*** & 12.08 & 12.71 & 0.71 &  \\ 
$\text{Regime duration}^2$ & 278.00 & 2783.57 & 0.00** & 278.00 & 287.38 & 0.89 &  \\ 
Military & 0.08 & 0.06 & 0.67 & 0.08 & 0.12 & 0.57 &  \\ 
Monarchy & 0.04 & 0.25 & 0.00*** & 0.04 & 0.04 & 1.00 &  \\ 
Party & 0.29 & 0.48 & 0.07* & 0.29 & 0.29 & 1.00 &  \\ 
log(gdp) & 22.65 & 23.95 & 0.00*** & 22.65 & 22.78 & 0.67 &  \\ 
log(gdp per capita) & 7.63 & 8.58 & 0.00*** & 7.63 & 7.65 & 0.95 &  \\ 
Military expenditure (\%GDP) & 2.44 & 4.34 & 0.00*** & 2.44 & 2.40 & 0.89 &  \\
Land (1000 $\text{km}^2$)& 760.22 & 1674.68 & 0.00*** & 760.21 & 783.37 & 0.80 &  \\ 
Population (million) & 22.62 & 140.73 & 0.00*** & 22.62 & 21.74 & 0.90 &  \\
Resource Export (\%GDP) & 14.79 & 22.64 & 0.03** & 14.79 & 14.16 & 0.72 &  \\ 
Resource $\times$ Ethnic & 8.61 & 14.43 & 0.01** & 8.61 & 8.33 & 0.82 & \\
  \hline
\end{tabular}
\caption{Pre and Post-Matching Balance}
\label{tab:matching_balance_largen}
\end{sidewaystable}

### 2014-11-22 writeups/multilevel_model.tex
done with all, doing references

Given that my dependent variable of interest (\textit{violence}) is at the event level, while my ``treatment'' (\textit{legislature}) is at the country year level, a Bayesian hierarchical model as the most appropriate choice.\footnote{The choice of Bayesian hierarchical model is largely practical: with 15,037 events, popular R packages to fit frequentist hierarchical model do not converge due to a high chance of numerical issues.}  Each event belongs to a country-year, and each country-year belongs to a country. The intercept at each level varies according to the group the unit belongs to. This model choice is superior to the alternative option, which is to aggregate the event-level data up to the country-year level and run a fixed effect model. Doing so will 1) ignore the variation at the event level, resulting in standard errors that are too small; 2) leave the aggregation mechanisms up to the researcher with no clear theoretical reasoning.\footnote{Should we measure the level of violence within a country year by averaging across events? Why not the median? Or count the number of positive / negative events? The choice is not likely to be theoretically driven and open to abuse.}

@@

I fit this model with a Gibbs sampler written in JAGS, running 3 chains, each has 11,000 iterations with 1,000 burn-ins. The mixing is excellent and the result is consistent across 3 chains. (See \autopageref{sec:modeldiagnostic} for more model diagnostics).

>>>>

Given that my dependent variable of interest (\textit{violence}) is at the event level, while my ``treatment'' (\textit{legislature}) is at the country year level, a Bayesian hierarchical model as the most appropriate choice.\footnote{The choice of Bayesian hierarchical model is largely practical: with 15,037 events, popular R packages to fit frequentist hierarchical model do not converge due to a high chance of numerical issues.}  Each event belongs to a country-year, and each country-year belongs to a country. The intercept at each level varies according to the group the unit belongs to. This model choice is superior to the alternative option, which is to aggregate the event-level data up to the country-year level and run a fixed effect model. Doing so will 1) ignore the variation at the event level, resulting in standard errors that are too small and thus false positives; 2) leave the aggregation mechanisms up to the researcher with no clear theoretical reasoning.\footnote{Should we measure the level of violence within a country year by averaging across events? Why not the median? Or count the number of positive / negative events? The choice is not likely to be theoretically driven and open to abuse.}

@@

I fit this model with a Gibbs sampler written in JAGS, running three chains, each has 11,000 iterations with 1,000 burn-ins. The mixing in the traceplot is excellent (i.e. having little autocorrelation) and does not vary across three chains, meaning that our sampling process has accurately approximated the true posterior distribution of the model's parameters. (See \autopageref{sec:modeldiagnostic} for more model diagnostics).

### 2014-11-22 writeups/multilevel_result.tex
done with all, doing references

From the posterior distribution, we know that the probability of legislature have$Pr(liec > 0) = 0.89$ close to significant.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & mean & 2.5\% & 97.5\% & Pr( $>$ 0) & Pr( $<$ 0) \\ 
  \hline
legislature & 0.48 & -0.29 & 1.25 & 0.89 &  \\ 
  log(GDP percap) & -0.39 & -1.31 & 0.52 &  & 0.81 \\ 
  log(GDP) & 0.37 & -0.19 & 0.92 & 0.91 &  \\ 
  resource (\%GDP) & -0.01 & -0.04 & 0.02 &  & 0.78 \\ 
  military exp (\%GDP) & -0.04 & -0.27 & 0.19 &  & 0.63 \\ 
  regime duration & 0.02 & -0.03 & 0.08 & 0.78 &  \\ 
  military regime & 2.10 & -1.79 & 5.84 & 0.86 &  \\ 
  personal regime & 1.60 & -1.65 & 4.89 & 0.83 &  \\ 
  party regime & 3.09 & -0.10 & 6.31 & 0.97 &  \\ 
   \hline
\end{tabular}
\caption{Hierarchical model parameter estimates}
\label{tab:hierarchical_estimate}
\end{table}

>>>>

Having obtained the posterior distribution of the parameters, we can calculate their means, 95\% credible interval, and the probabilities of them being different from 0 (See \autoref{tab:hierarchical_estimate}). Having an legislature improves the positivity in regime-dissident interaction by 0.48 point along the Goldstein scale, and we are 89\% certain that this effect is truly positive. In addition, a party dictatorship improves the interaction by 3.09, a substantively significant effect.

While this model result provides some support for the moderation effect of authoritarian legislature on political violence, the size of the substantive effect is quite small given that regime-dissdent interactions typically scatter along a 10-unit spectrum on the Goldstein scale (-10 to 0). Since our model allows full flexibility in the country-year and country intercepts, it reveals that much of the variation in violence is captured in these group-specific intercepts. This suggests that the included set of co-variates, which is standard in the literature, have not fully explained the pattern of violence.

\begin{table}[H]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & mean & 2.5\% & 97.5\% & Pr( $>$ 0) & Pr( $<$ 0) \\ 
  \hline
legislature & 0.48 & -0.29 & 1.25 & 0.89 &  \\ 
  log(GDP percap) & -0.39 & -1.31 & 0.52 &  & 0.81 \\ 
  log(GDP) & 0.37 & -0.19 & 0.92 & 0.91 &  \\ 
  resource (\%GDP) & -0.01 & -0.04 & 0.02 &  & 0.78 \\ 
  military exp (\%GDP) & -0.04 & -0.27 & 0.19 &  & 0.63 \\ 
  regime duration & 0.02 & -0.03 & 0.08 & 0.78 &  \\ 
  military regime & 2.10 & -1.79 & 5.84 & 0.86 &  \\ 
  personal regime & 1.60 & -1.65 & 4.89 & 0.83 &  \\ 
  party regime & 3.09 & -0.10 & 6.31 & 0.97 &  \\ 
   \hline
\end{tabular}
\caption{Hierarchical model parameter estimates}
\label{tab:hierarchical_estimate}
\end{table}

### 2014-11-22 writeups/theory.tex
done with all, doing references

\section{Theory: Authoritarian legislature and its effect on political violence}

@@

To consider whether violent means is chosen to reach their goal, we need a theoretical framework to understand how the opposition and the regime engage in the bargaining game. A possible starting point is the bargaining models developed for democratic legislatures, where bargaining outcomes are driven by various institutional rules (positive and negative rule, sequencing of these powers, first-mover advantage, etc.) (Baron and Ferejohn 1989; Cox 2006). However, this type model is not suited to analyze the bargaining under an authoritarian regime, where the institutional rules themselves are under constant threat of being revised by the final arbiter that is violence. In other words, the end game of legislative bargaining now has an extra node that is whether either side will violently suppress the other’s demand. Given this context, the decision of the autocrat and the opposition is better characterized as a bargaining game in anarchy, very much like the bargaining between two sovereign states with the option of going to war (Fearon 1995).

To formally model the interaction, let the negotiation between the opposition and the regime be about the share $x$, $(0 < x < 1)$ of a good (normalized as 1) given to the opposition. (For concreteness we can think of this as the share of the economy’s rent or capital stock). Both sides have the option 1) to bargain and decide on x, or 2) to engage in violent protest / repression. Violence is costly compared to peaceful negotiation---we call this difference in relative cost $c_A$ and $c_O$ for the autocrat and the opposition, respectively. If violence happens, whichever side emerges victorious decides the share x to his heart’s desire---i.e. the winning autocrat would give 0 to the opposition and the winning opposition would give 1 to itself.

Given the uncertain nature of collective action battles, the utility of engaging in violence for the two sides is best formulated as an expected utility that depends on the probability of the opposition winning (call this probability $p$). Thus, the expected utility of the opposition is $p – c_O$, while the expected utility of the ruling regime is $1 - p - c_A$. These two expected utilities sum to less than one, which raises the puzzle why the two sides fail to engage in ex-ante bargaining that splits the difference, giving both sides a higher payoff than the expected utility of violence. Indeed, any amount of x between $(p - c_O, p + c_A)$, i.e. the bargaining range, is more desirable than violence to both sides.

@@

\item The bargaining range $(p - c_O, p + c_A)$ depends on the value of $c_O and c_A$

@@

H1: Authoritarian regimes that create a legislature will experience less violence in its interaction with the dissidents.
\end{enumerate}

\subsection{Legislature increasing violence}
Throughout the previous model, it is implied that the actors’ utility functions remain static with the only parameters being the cost of violence and the share of the issue space. Yet this assumption may miss the dynamism of the opposition’s democratic expectation, which becomes increasingly demanding as the regime starts liberalizing.

Speaking about this issue, Ted Gurr hypothesizes, “The potential for collective violence varies strongly with the intensity and the scope of relative deprivation among members of a collectivity.” Relative deprivation, in turn, is defined as the perceived discrepancy between value expectation (what people believe they are entitled to) and value capability (what the people believe they are able to attain) (1970, 24). Seen under this light, repeated elections may increase the potential for political violence by raising the democratic expectation, which can lead to violent protest if unfulfilled. The effect of elections is especially potent because it affects all citizens, thus raising not only the intensity but also the scope of relative deprivation.

Indeed, unfulfilled political expectation is a common case of relative deprivation causing violence. Historical examples are numerous, spanning various time periods as well as scopes of conflict: from the Puritan Revolution of 1640-60, which was sparked by attempts of the Stuart kings to reinstate royal absolutism, to the protest in twentieth-century colonies, which often followed the imposition of restrictions after a period of political liberalization (1970, 115). This pattern makes sense because, when people have tasted freedom, not only do they want more rights but they also consider rights more important. In other words, the opposition may not simply optimize its share of the issue space but also develops a minimum level of acceptable concession. If opposition’s expectation rises too quickly, the ruling regime may not have enough time to appropriate adjust their position especially if the hard-liners are also trying to retrench given signs of a political spiral. For this reason, it is plausible that electoral authoritarianism leads to more violent interaction between the ruling regime and the opposition.

>>>>

\section{Theory: Authoritarian legislature and its effect on political violence}
\label{sec:theory}

@@

To consider whether violent means is chosen to reach their goal, we need a theoretical framework to understand how the opposition and the regime engage in the bargaining game. A possible starting point is the bargaining models developed for democratic legislatures, where bargaining outcomes are driven by various institutional rules (positive and negative rule, sequencing of these powers, first-mover advantage, etc.) (Baron and Ferejohn 1989; Cox 2006). However, this type model is not suited to analyze the bargaining under an authoritarian regime, where the institutional rules themselves are under constant threat of being revised by the final arbiter that is violence. In other words, the end game of legislative bargaining now has an extra node: whether either side will violently suppress the other’s demand. Given this context, the decision of the autocrat and the opposition is better characterized as a bargaining game in anarchy, very much like the bargaining between two sovereign states with the option of going to war (Fearon 1995).

To formally model the interaction, let the negotiation between the opposition and the regime be about the share $x$, $(0 < x < 1)$ of a good (normalized as 1) given to the opposition.\footnote{For concreteness we can think of this as the share of the economy’s rent or capital stock}. Both sides have the option 1) to bargain and decide on x, or 2) to engage in violent protest / repression. Violence is costly compared to peaceful negotiation---we call this difference in relative cost $c_A$ and $c_O$ for the autocrat and the opposition, respectively. If violence happens, whichever side emerges victorious decides the share x to his heart’s desire---i.e. the winning autocrat would give 0 to the opposition and the winning opposition would give 1 to itself.

Given the uncertain nature of collective action battles, the utility of engaging in violence for the two sides is best formulated as an expected utility that depends on the probability of the opposition winning (call this probability $p$). Thus, the expected utility of the opposition is $p – c_O$, while the expected utility of the ruling regime is $1 - p - c_A$. These two expected utilities sum to less than one ($1 - c_A - c_O < 1$), which raises the puzzle why two sides fail to engage in ex-ante bargaining that splits the difference, giving both sides a higher payoff than the expected utility of violence. Indeed, any amount of x between $(p - c_O, p + c_A)$, i.e. the bargaining range, is more desirable than violence to both sides.

@@

\item The bargaining range $(p - c_O, p + c_A)$ depends on the value of $c_O$ and $c_A$

@@

\begin{quote}
H1: Authoritarian regimes that create a legislature will experience less violence in their interaction with the dissidents.
\end{quote}
\end{enumerate}

\subsection{Legislature increasing violence}
While the prevalent theory of legislature as an informational and commitment device favors hypothesis H1, it is important to note that there are reasons to susepct the opposite. Throughout the previous model, it is implied that the actors’ utility functions remain static with the only parameters being the cost of violence and the share of the issue space. Yet this assumption may miss the dynamism of the opposition’s democratic expectation, which becomes increasingly demanding as the regime starts liberalizing.

Speaking about this issue, Ted Gurr hypothesizes, ``The potential for collective violence varies strongly with the intensity and the scope of relative deprivation among members of a collectivity.'' Relative deprivation, in turn, is defined as the perceived discrepancy between value expectation (what people believe they are entitled to) and value capability (what the people believe they are able to attain) (1970, 24). Seen under this light, repeated elections may increase the potential for political violence by raising the democratic expectation, which can lead to violent protest if unfulfilled. The effect of elections is especially potent because it affects all citizens, thus raising not only the intensity but also the scope of relative deprivation.

Indeed, unfulfilled political expectation is a common case of relative deprivation causing violence. Historical examples are numerous, spanning various time periods as well as scopes of conflict: from the Puritan Revolution of 1640-60, which was sparked by attempts of the Stuart kings to reinstate royal absolutism, to the protest in twentieth-century colonies, which often followed the imposition of restrictions after a period of political liberalization (1970, 115). This pattern makes sense because, when people have tasted freedom, not only do they want more rights but they also consider rights more important. In other words, the opposition may not simply optimize its share of the issue space but also develops a minimum level of acceptable concession. If opposition’s expectation rises too quickly, the ruling regime may not have enough time to appropriate adjust their position especially if the hard-liners are also trying to retrench given signs of a political spiral. For this reason, it is plausible that electoral authoritarianism leads to more violent interaction between the ruling regime and the opposition, leading to the second hypothesis:

\begin{quote}
H2: Authoritarian regimes that create a legislature will experience more violence in their interaction with the dissidents.
\end{quote}
